---
title: "hw2"
author: "ShibeiXu"
date: "2021/10/7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r,message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```
### load the trash data with data cleaning
```{r load trash data,message=FALSE,warning=FALSE}
trash_data=read_excel("./data/trash_data.xlsx",sheet = 1,range = cell_cols(1:14))%>%
  janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls,digits = 0))
head(trash_data,5)
```
### Load the 2018 and 2019 data and combine into one table
```{r load 2018 2019 precipitation data,message=FALSE,warning=FALSE}
prep_2018_data=read_excel("./data/trash_data.xlsx",sheet = 5,skip=1,col_names = TRUE)%>%
  janitor::clean_names()%>%
  drop_na(month)%>%
  mutate(year=2018)
prep_2019_data=read_excel("./data/trash_data.xlsx",sheet = 4,skip=1,col_names = TRUE)%>%
  janitor::clean_names()%>%
  drop_na(month)%>%
  mutate(year=2019)
head(prep_2018_data,5)
head(prep_2019_data,5)
```

```{r combine 2018 and 2019,message=FALSE,warning=FALSE}
total_table=bind_rows(prep_2018_data,prep_2019_data)%>%
  mutate(month=month.name[as.numeric(month)])
head(total_table,5)
```

### Analysis
The combined 2018 and 2019 precipitation data is a `r size_sum(total_table)` dataset with `r ncol(total_table)` variables and `r nrow(total_table)` data input. The variables of this dataset includes `r names(total_table)`. The total precipitation in 2018 is `r sum(pull(prep_2018_data, total),na.rm=TRUE)`.

The trash table is a `r size_sum(trash_data)` dataser with `r ncol(trash_data)` variable and `r nrow(trash_data)` observations. The variables includes `r names(trash_data)`. The median number of sports balls in a dumpster in 2019 is `r trash_data %>% filter(year==2019)%>%pull(sports_balls)%>%median()` `

## Problem 2
### clean the data in pols_month
```{r clean pols-month data,message=FALSE,warning=FALSE}
poll_data=read_csv("./data/pols-month.csv")%>%separate("mon",into=c("year","month","date"))%>%mutate(month=month.name[as.numeric(month)],month=str_to_lower(month),president=recode(prez_gop,"0"="dem","1"="gop"))%>%janitor::clean_names()
poll_data=select(poll_data,-date,-prez_dem,-prez_gop)
head(poll_data,5)
```
```{r load snp, message=FALSE,warning=FALSE}
snp=read_csv("data/snp.csv")%>%separate("date",into=c("year","month","day"),sep = "-")%>%mutate(month=month.name[as.numeric(month)],month=str_to_lower(month))%>%janitor::clean_names()
snp=select(snp,-day)
head(snp,5)
```
```{r load and clean the unemployment date, message=FALSE, warning=FALSE}
unemploy=read.csv("./data/unemployment.csv")%>%
  janitor::clean_names()%>%
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment rate")%>%
  mutate(year=as.character(year),
    month=recode(month,jan="january",feb="february",mar="march",apr="april",jun="june",jul="july",aug="august",sep="september",oct="october",nov="november",dec="december"))
head(unemploy,5)
```

```{r bind the table,warning=FALSE,message=FALSE}
poll_1_data=left_join(poll_data,snp)
poll_full=left_join(poll_1_data,unemploy)
head(poll_full,5)
```
The combined poll,snp and unemployment data is a `r size_sum(poll_full)` dataset with `r ncol(poll_full)` variables and `r nrow(poll_full)` data input. The variables of this dataset includes `r names(poll_full)`. The total range of the year included in this data is `r poll_full%>% pull(year)%>%range()`.

## Problem 3
```{r load the baby names dataset and tidy the data,message=FALSE,warning=FALSE}
baby_data=read_csv("./data/baby_name.csv")%>%
  janitor::clean_names()%>%
  mutate(
    childs_first_name=str_to_lower(childs_first_name),
    ethnicity=str_to_lower(ethnicity),
    gender=str_to_lower(gender),
    ethnicity=recode(ethnicity,"asian and paci"="asian and pacific islander","black non hisp"="black non hispanic","white non hisp"="white non hispanic"))%>%
    distinct()
head(baby_data,5)
```
### produce the table for the rank of the name olivia
```{r table for the name Olivia,warning=FALSE,message=FALSE}
olivia_table=filter(baby_data,childs_first_name=="olivia")%>%
  select(-gender,-childs_first_name,-count)%>%
  mutate(year_of_birth=as.factor(year_of_birth))
head(olivia_table,5)
```

### pivot the year into columns
```{r}
pivot_wider(olivia_table,names_from = year_of_birth,values_from = rank)
```
### find the most popular name over time among male children
```{r,warning=FALSE,message=FALSE}
male_table=filter(baby_data,rank==1&gender=="male")%>%
select(-gender,-count,-rank)%>%
  mutate(year_of_birth=as.factor(year_of_birth))
head(male_table,5)
```
### Pivot the year into columns
```{r} 
pivot_wider(male_table,names_from = year_of_birth,values_from = childs_first_name)
```
### produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis) for male, white non-hispanic children born in 2016
```{r filter the male,white non-hispanic children born in 2016}
graph_data=filter(baby_data,gender=="male"&year_of_birth==2016&ethnicity=="white non hispanic")
```
```{r produce the scatterplot}
ggplot(graph_data,aes(x=rank,y=count))+geom_point()
```


